---
title: "Final Project First Draft"
output: pdf_document
---
\maketitle

#### Running Title:
07 - 405 - Crimes in Chicago

#### Group Members:
Benjamin Mao, Cecilia Xin, Monica Liu, Jared Boyd

#### Introduction to Project and Dataset:
Our goal for this project is to find patterns in crime in Chicago. Our dataset
is sourced from the Chicago Police Department's CLEAR system, comprising 
reported incidents of crime in the City of Chicago from 2001 to the present, 
with addresses anonymized at the block level for victim privacy. This data
includes the date, time, type, and general location of crime in Chicago.

#### Load in the Data
---

```{r}
#library(readr)
#crimes_data <- read_csv("Crimes_-_2001_to_Present.csv")
```

## Histogram
With this plot we will analyze the number of crimes in different years

```{r}

```

## Curve
With this plot we will analyze the number of crimes within a year/day

```{r}

```

## Pie Chart
With this plot we will analyze the frequency of the different crime types

```{r}

```

## Barplot
With this plot we will analyze the location of the crimes

```{r}
# Plot the location of the crimes using a barplot
barplot(table(crimes_data$LocationDescription), main="Location of Crimes", xlab="Location", ylab="Frequency")

```

## Boxplot

```{r}

```

## Scatterplot

```{r}

```
