---
title: "Final Project Second Draft"
output: pdf_document
---
\maketitle

#### Running Title:
07 - 405 - Crimes in Chicago

#### Group Members:
Benjamin Mao, Cecilia Xin, Monica Liu, Jared Boyd

#### Introduction to Project and Dataset:
Our goal for this project is to find patterns in crime in Chicago. Our dataset
is sourced from the Chicago Police Department's CLEAR system, comprising 
reported incidents of crime in the City of Chicago from 2001 to the present, 
with addresses anonymized at the block level for victim privacy. This data
includes the date, time, type, and general location of crime in Chicago.

#### Load in the Data
---

```{r}
library(readr)
crimes_data <- read_csv("Crimes_-_2001_to_Present.csv")
```

#### Six Standard Plots
## Histogram
With this plot we will analyze the number of crimes in different years.
The graph shows that the number of crimes has been decreasing over time.

```{r}
hist(crimes_data$Year, main = "Distribution of Years", xlab = "Year", 
     ylab = "Frequency", col = "skyblue", border = "white")
```

## Curve
With this plot we will analyze arrest rate over time.
The graph shows that the arrest rate has been decreasing over time.


```{r}
# Plot the year on the x-axis and the arrest rate (calculated per year) on the y-axis to see if arrest rates are increasing or decreasing over time. In the Arrest column, True means an arrest was made, and False means no arrest was made.

# install.packages("dplyr")
library(dplyr)
arrests_by_year <- crimes_data %>% 
  group_by(Year) %>% 
  summarise(Arrest_Rate = sum(Arrest == "TRUE") / n())
# arrests_by_year
```


```{r}
# curve

plot(arrests_by_year$Year, arrests_by_year$Arrest_Rate, 
     type = "o",  
     xlab = "Year", ylab = "Arrest Rate",
     main = "Arrest Rate from 2001 to 2023",
     col = "pink", pch = 16, ylim = c(0, 1))
```


## Pie Chart
With this plot we will analyze the frequency of the different crime types.
The graph shows that theft is the most common crime type, followed by battery and criminal damage.

```{r}
crime_type <- table(crimes_data$`Primary Type`)
total_crimes <- sum(crime_type)
crime_type_percent <- prop.table(crime_type) * 100

# sort types with frequencies below 3% into "Other"category 
other_crime_types <- names(crime_type_percent)[crime_type_percent < 3]
crime_type["Other"] <- sum(crime_type[other_crime_types])
crime_type <- crime_type[!(names(crime_type) %in% other_crime_types)]

# plot Primary Type pie chart
pie(crime_type, main = "Distribution of Crimes by Type", col = rainbow(length(crime_type)))

# Also plot the percentage in another graph
par(xpd = TRUE)
plot.new()

legend("topright", legend = paste(names(crime_type), "(", round(crime_type_percent[1:length(crime_type)], 2), "%)"), cex = 0.8, fill = rainbow(length(crime_type)))


```

## Barplot
With this plot we will analyze the location of the crimes.
The graph shows that most crimes occur in the street, followed by residences and apartments.

```{r}
# Plot the location of the crimes using a barplot
table = table(crimes_data$`Location Description`)
ordered = table[order(table, decreasing = TRUE)][1:5]

# Make the x-axis labels more readable
names(ordered) = c("Street", "Residence", "Apartment", "Sidewalk", "Other")

# Plot bar plot
barplot(ordered, main = "Top 5 Locations of Crimes", xlab = "Location", ylab = "Frequency", col = "lightblue")




```

## Scatter Plot
With this plot we will analyze the relationship between the latitude and longitude of the crimes.
The graph shows that most crimes occur in the city center, with a few outliers in the suburbs.
Future improvement: use ggplot to create a more detailed and visually appealing scatterplot.

```{r}
# Print a scatterplot of lattitude versus longitude. This could reveal if certain areas of the city are more prone to crime. Don't include ggplot.
sampled_crimes <- crimes_data[sample(1:nrow(crimes_data), 10000), ]

plot(sampled_crimes$Longitude, sampled_crimes$Latitude, 
     xlab = "Longitude", ylab = "Latitude",
     main = "Scatterplot of Crime Locations",
     xlim = c(-87.85, -87.5), ylim = c(41.6, 41.9), pch = 20, col = "blue", cex = 0.5, bg = "white")

```

## Density Plot
This is a density plot showing how crime incidents are distributed across different times of the day, with the x-axis representing hours since midnight and the y-axis representing the density of crime occurrences.
The graph shows that crime is most likely to occur between 12:00 PM and 12:00 AM, with a peak around 8:00 PM.

```{r}
# Assuming your dataset is named 'data' and the date column is named 'Date'

# Step 1: Convert 'Date' to POSIXct (date-time object) and extract time
crimes_data$Time <- format(strptime(crimes_data$Date, format = "%m/%d/%Y %I:%M:%S %p"), format = "%H:%M:%S")

# Step 2: Convert Time to hours since midnight
crimes_data$Hours <- as.numeric(substr(crimes_data$Time, 1, 2)) + 
              as.numeric(substr(crimes_data$Time, 4, 5))/60 +
              as.numeric(substr(crimes_data$Time, 7, 8))/3600

# Step 3: Create the density plot
plot(density(crimes_data$Hours, na.rm = TRUE), 
     main = "Density of Crimes Over Time of Day",
     xlab = "Hours (since midnight)",
     ylab = "Density")


```

#### Statistical modeling and analysis
We will use a linear regression model to analyze the relationship between 
the number of cannabis possession crimes and the number of heroin 
possession crimes per year.

As seen in the graph below, there is a pretty strong relation between
the crimes of these two drugs in Chicago. Further in this project we intend
to look more into this relation and try and find others.
```{r}
cannabis = c()
heroin = c()

for(year in 2001:2024){
  which = crimes_data$Year == year
  desc = crimes_data$Description[which]
  total_c = sum(grepl("CANNABIS", desc))
  total_h = sum(grepl("HEROIN", desc))
  cannabis = c(cannabis, total_c)
  heroin = c(heroin, total_h)
}
reg_line = lm(heroin ~ cannabis)

plot(x = cannabis, y = heroin, 
     main="Linear Regression of Heroin Use on Cannabis Use", 
     col = "blue", xlab="Amount of Cannabis Crimes per Year",
     ylab="Amount of Heroin Crimes per Year")
abline(reg_line[[1]][1], reg_line[[1]][2], col = "red")
```

#### Six GGPlots

## Graph 1
```{r}

```

## Graph 2
```{r}

```

## Graph 3
```{r}

```

## Graph 4
```{r}

```

## Graph 5
```{r}

```

## Graph 6
```{r}

```


